The role of cross-linguistic lexical similarity on bilingual word acquisition

### 2021-09-30 Rmd/manuscript.tex
Fix comprehension model and update report

Despite this variability, children's trajectories of vocabulary growth seem quite stable across languages. Tardif et al. (2008) collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing. Since birth, these infants had been learning English, Cantonese and Mandarin, respectively. The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (\emph{daddy}, \emph{mommy}), social routines (\emph{bye}, \emph{uh-oh}) or animals (\emph{woof-woof}). These results were later extended by Frank, Braginsky, Yurovsky, and Marchman (2021) to a wider diversity languages, who also reported that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages. Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.

@@

Floccia et al.~pointed to \emph{parallel activation} as the main mechanism underpinning their results. The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension. This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production (Costa, Caramazza, \& Sebastian-Galles, 2000; Hoshino \& Kroll, 2008) or comprehension (Thierry \& Wu, 2007). One of the clearest examples of parallel activation was provided by Costa, Caramazza, and Sebastian-Galles (2000). In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish. In half of the trials, the object labels were cognates in Spanish and Catalan (\emph{árbol}-\emph{arbre}, translations of \emph{tree}), whereas in the other half of the trials labels were non-cognates (\emph{mesa}-\emph{taula}, translations of \emph{table}). Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items. Importantly, Spanish monolinguals did not show this effect. These results suggest that, for the bilinguals, Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures. Several recent studies have also provided similar evidence in comprehension in children (Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013; e.g., Von Holzen \& Mani, 2012). Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes. A missing piece in this account is how increase parallel activation might facilitate the acquisition new words.

@@

\begin{itemize}
\tightlist
\item
  \emph{Age} of participant in months (\texttt{age}), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
\item
  \emph{Lexical frequency} of item (\texttt{frequency\_center}), Relative lexical frequencies (frequency per million words, \(fpmw\)) were extracted from the CHILDES data base, using the \texttt{childesr} R package (M. Braginsky, Sanchez, \& Yurovsky, 2018), and were transformed to Zipf scores (Van Heuven, Mandera, Keuleers, \& Brysbaert, 2014; Zipf, 1949) (\(log10(fpmw)+3\)). Missing frequencies were imputed by multiple imputation using the \texttt{mice} R package {[}van2011mice{]}. The resulting scores were then centred around the mean.
\item
  \emph{Degree of exposure} to the language (\texttt{doe\_center}), computed as the proportion of the time the child is exposed to the language the item belongs to. For instance, for a response to the item \texttt{perro} (Spanish), the degree of exposure would correspond to the participant's relative exposure to Spanish (ranging from 0 to 1). Before entering the analyses, this variable was centred it around the mean and divided by 10, so that the values of the regression coefficients of this predictor correspond to a 10\% increase in exposure to the language of the item.
\item
  \emph{Cognateness}, a categorical variable indicating whether the Catalan and Spanish forms of the TE are identical cognates (IC), non-identical cognates (NIC), or non-cognates (NC), as rated by a trained Spanish-Catalan bilingual linguist. We sum-coded two contrasts (Schad, Vasishth, Hohenstein, \& Kliegl, 2020): \texttt{cognate1}, comparing IC and NIC (both coded as +0.25) with NC (coded as -0.5) and \texttt{cognate2}, comparing IC (coded as +0.5) with NIC (coded as -0.5).
\end{itemize}

We first fit a base model that only included the main effects of \texttt{age} and \texttt{frequency\_center} (Model 1) (see Mika Braginsky, Yurovsky, Marchman, \& Frank, 2019; Jones \& Brandt, 2019; Kachergis, Marchman, \& Frank, 2021 for similar approaches). We then entered our predictors of interest by adding \texttt{doe\_center} (Model 2) and then \texttt{cognate} and its interaction with \texttt{doe\_center} (Model 3). To account for the hierarchical structure of the data (some responses to different TEs corresponded to the same participant, and some responses from different participants corresponded to the same TE )every model included \emph{Participants} (\texttt{id}) and \emph{TEs} (\texttt{te}) as grouping variables (i.e.~random effects), with all possible random intercepts and slopes where appropriate (Barr, Levy, Scheepers, \& Tily, 2013) (see Appendix for more details of the model).

@@

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{lrrrrrrrrr}
\toprule
& \multicolumn{3}{c}{\textbf{LOOELPD}} & \multicolumn{3}{c}{\textbf{LOOIC}} & \multicolumn{3}{c}{\textbf{LOODiff.}} \\ 
 \cmidrule(lr){2-4}\cmidrule(lr){5-7}\cmidrule(lr){8-10}
Model & \emph{ELPD} & \emph{SE} & \emph{SE} (sub) & \emph{LOO-IC} & \emph{SE} & \emph{SE} (sub) & \emph{diff} & \emph{SE} & \emph{SE} (sub) \\ 
\midrule
\multicolumn{1}{l}{Comprehension} \\ 
\midrule
Model 1 & $-83,789.09$ & $211.11$ & $114.66$ & $167,578.18$ & $422.21$ & $229.32$ & $0.00$ & $0.00$ & $0.00$ \\ 
Model 2 & $-233,348.92$ & $682.60$ & $3,067.85$ & $466,697.84$ & $1,365.20$ & $6,135.70$ & $149,559.83$ & $714.50$ & $3,069.99$ \\ 
Model 3 & $-242,729.62$ & $721.85$ & $3,523.44$ & $485,459.24$ & $1,443.70$ & $7,046.88$ & $158,940.53$ & $752.08$ & $3,525.30$ \\ 
Model 1 & $-59,417.81$ & $209.84$ & $103.31$ & $118,835.62$ & $419.69$ & $206.61$ & $0.00$ & $0.00$ & $0.00$ \\ 
Model 2 & $-60,421.29$ & $206.95$ & $80.42$ & $120,842.58$ & $413.91$ & $160.85$ & $1,003.48$ & $294.73$ & $130.92$ \\ 
Model 3 & $-60,481.42$ & $208.06$ & $105.51$ & $120,962.85$ & $416.12$ & $211.02$ & $1,063.61$ & $295.50$ & $147.67$ \\ 
\bottomrule
\end{longtable}

@@

The average probability of participants understanding a TE was 53.47\% (intercept) when all predictors were equal to 0: mean participant age (0.00 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average acros the three levels). An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 2.93\% increment in comprehension probability. A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a 7.09\% increment in comprehension probability. A 10\% increase in relative language exposure made TEs -3.00\% more likely to be understood. Cognates (identical and non-identical) were 1.95\% more like to be understood than non-cognates. The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0. Identical cognates were 2.06\% less like to be understood than non-identical cognates. The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero). Figure 1 summarises the posterior distribution of the fixed effects of the model.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{llllrrr}
\toprule
& \multicolumn{3}{c}{Comprehension} & \multicolumn{3}{c}{Production} \\ 
 \cmidrule(lr){2-4}\cmidrule(lr){5-7}
\textbf{Predictor} & \textbf{Mean} & \textbf{SE} & \textbf{95\% CrI} & \textbf{Mean} & \textbf{SE} & \textbf{95\% CrI} \\ 
\midrule
Intercept\textsuperscript{1} & $53.47\%$ & $51.30\%$ & [$51.07\%$, $55.99\%$] & $16.24\%$ & $51.39\%$ & [$14.92\%$, $17.82\%$] \\ 
age\textsuperscript{2} & $2.93\%$ & $0.12\%$ & [$2.69\%$, $3.15\%$] & $5.57\%$ & $0.14\%$ & [$5.27\%$, $5.83\%$] \\ 
frequency\_center & $7.09\%$ & $0.51\%$ & [$6.09\%$, $8.00\%$] & $7.57\%$ & $0.61\%$ & [$6.34\%$, $8.77\%$] \\ 
doe\_center\textsuperscript{2} & $-3.00\%$ & $0.56\%$ & [$-4.12\%$, $-1.92\%$] & $-0.67\%$ & $0.57\%$ & [$-1.74\%$, $0.41\%$] \\ 
cognate1\textsuperscript{2} & $1.95\%$ & $2.19\%$ & [$-2.33\%$, $6.16\%$] & $2.91\%$ & $2.01\%$ & [$-1.15\%$, $6.87\%$] \\ 
cognate2 & $-2.06\%$ & $1.85\%$ & [$-5.71\%$, $1.49\%$] & $-0.46\%$ & $2.09\%$ & [$-4.88\%$, $3.30\%$] \\ 
doe\_center:cognate1\textsuperscript{2} & $0.60\%$ & $0.72\%$ & [$-0.72\%$, $2.13\%$] & $-0.72\%$ & $0.70\%$ & [$-2.05\%$, $0.62\%$] \\ 
doe\_center:cognate2 & $1.76\%$ & $0.68\%$ & [$0.47\%$, $3.10\%$] & $0.60\%$ & $0.72\%$ & [$-0.82\%$, $2.01\%$] \\ 
\bottomrule
\end{longtable}
\vspace{-5mm}
\begin{minipage}{\linewidth}
\textsuperscript{1}Transformed using the inverse logit to get the average probability of correct response \\ 
\textsuperscript{2}Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable. \\ 
\end{minipage}

@@

The average probability of participants understanding a TE was 16.24\% (intercept) when all predictors were equal to 0: mean participant age (0.00 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average acros the three levels). An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 5.57\% increment in comprehension probability. A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a 7.57\% increment in comprehension probability. A 10\% increase in relative language exposure made TEs -0.67\% more likely to be understood. Cognates (identical and non-identical) were 2.91\% more like to be understood than non-cognates. The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0. Identical cognates were 2.06\% less like to be understood than non-identical cognates. The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero). Figure 1 summarises the posterior distribution of the fixed effects of the model.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{lcccrrr}
\toprule
\textbf{Predictor} & \textbf{Mean} & \textbf{SEM} & \textbf{95\% CrI} & \textbf{Rhat} & \textbf{Bulk ESS}\textsuperscript{1} & \textbf{Tail ESS}\textsuperscript{1} \\ 
\midrule
Intercept\textsuperscript{2} & $16.24\%$ & $51.39\%$ & [$14.92\%$, $17.82\%$] & $1.05$ & $102$ & $201$ \\ 
age\textsuperscript{3} & $5.57\%$ & $0.14\%$ & [$5.27\%$, $5.83\%$] & $1.01$ & $540$ & $504$ \\ 
frequency\_center & $7.57\%$ & $0.61\%$ & [$6.34\%$, $8.77\%$] & $1.00$ & $445$ & $432$ \\ 
doe\_center\textsuperscript{3} & $-0.67\%$ & $0.57\%$ & [$-1.74\%$, $0.41\%$] & $1.00$ & $847$ & $849$ \\ 
cognate1\textsuperscript{3} & $2.91\%$ & $2.01\%$ & [$-1.15\%$, $6.87\%$] & $1.01$ & $260$ & $518$ \\ 
cognate2 & $-0.46\%$ & $2.09\%$ & [$-4.88\%$, $3.30\%$] & $1.02$ & $203$ & $312$ \\ 
doe\_center:cognate1\textsuperscript{3} & $-0.72\%$ & $0.70\%$ & [$-2.05\%$, $0.62\%$] & $1.01$ & $1,119$ & $838$ \\ 
doe\_center:cognate2 & $0.60\%$ & $0.72\%$ & [$-0.82\%$, $2.01\%$] & $1.00$ & $1,279$ & $743$ \\ 
\bottomrule
\end{longtable}
\vspace{-5mm}
\begin{minipage}{\linewidth}
\textsuperscript{1}ESS: Effective sample size \\ 
\textsuperscript{2}Transformed using the inverse logit to get the average probability of correct response \\ 
\textsuperscript{3}Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable. \\ 
\end{minipage}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-agresti2003categorical}{}%
Agresti, A. (2003). \emph{Categorical data analysis} (Vol. 482). John Wiley \& Sons.

\leavevmode\hypertarget{ref-arslan2020formr}{}%
Arslan, R. C., Walther, M. P., \& Tata, C. S. (2020). Formr: A study framework allowing for automated feedback generation and complex longitudinal experience-sampling studies using r. \emph{Behavior Research Methods}, \emph{52}(1), 376--387.

\leavevmode\hypertarget{ref-barr2013random}{}%
Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random effects structure for confirmatory hypothesis testing: Keep it maximal. \emph{Journal of Memory and Language}, \emph{68}(3), 255--278.

\leavevmode\hypertarget{ref-bergelson2020comprehension}{}%
Bergelson, E. (2020). The comprehension boost in early word learning: Older infants are better learners. \emph{Child Development Perspectives}, \emph{14}(3), 142--149.

\leavevmode\hypertarget{ref-bergelson2012months}{}%
Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants know the meanings of many common nouns. \emph{Proceedings of the National Academy of Sciences}, \emph{109}(9), 3253--3258.

\leavevmode\hypertarget{ref-bergelson2015early}{}%
Bergelson, E., \& Swingley, D. (2015). Early word comprehension in infants: Replication and extension. \emph{Language Learning and Development}, \emph{11}(4), 369--380.

\leavevmode\hypertarget{ref-bloom2002children}{}%
Bloom, P. (2002). \emph{How children learn the meanings of words}. MIT press.

\leavevmode\hypertarget{ref-bosch2014first}{}%
Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter? \emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}

\leavevmode\hypertarget{ref-bosch2014first}{}%
Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter? \emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}

\leavevmode\hypertarget{ref-braginsky2018childesr}{}%
Braginsky, M., Sanchez, A., \& Yurovsky, D. (2018). Childesr: Accessing the `CHILDES'database. \emph{R Package Version 0.1. 0}.

\leavevmode\hypertarget{ref-braginsky2019consistency}{}%
Braginsky, Mika, Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019). Consistency and variability in children's word learning across languages. \emph{Open Mind}, \emph{3}, 52--67.

\leavevmode\hypertarget{ref-burkner2017brms}{}%
Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.

\leavevmode\hypertarget{ref-carpenter2017stan}{}%
Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A. (2017). Stan: A probabilistic programming language. \emph{Journal of Statistical Software}, \emph{76}(1), 1--32.

\leavevmode\hypertarget{ref-cattani2014much}{}%
Cattani, A., Abbot-Smith, K., Farag, R., Krott, A., Arreckx, F., Dennis, I., \& Floccia, C. (2014). How much exposure to english is necessary for a bilingual toddler to perform like a monolingual peer in language tests? \emph{International Journal of Language \& Communication Disorders}, \emph{49}(6), 649--671.

\leavevmode\hypertarget{ref-costa2000cognate}{}%
Costa, A., Caramazza, A., \& Sebastian-Galles, N. (2000). The cognate facilitation effect: Implications for models of lexical access. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{26}(5), 1283.

\leavevmode\hypertarget{ref-david2008individual}{}%
David, A., \& Wei, L. (2008). Individual differences in the lexical development of french--english bilingual children. \emph{International Journal of Bilingual Education and Bilingualism}, \emph{11}(5), 598--618.

>>>>

Despite this variability, children's trajectories of vocabulary growth seem quite stable across languages. Tardif et al. (2008) collected data about the first ten words acquired by 10 to 16 month-old infants living in the United States, Hong Kong, and Beijing. Since birth, these infants had been learning English, Cantonese and Mandarin, respectively. The authors found a common pattern across the three groups: their first ten words referred to roughly the same concepts, namely relatives/caretakers (\emph{daddy}, \emph{mommy}), social routines (\emph{bye}, \emph{uh-oh}) or animals (\emph{woof-woof}). These results were later extended by Frank et al. (2021) to a wider diversity languages, who also reported that such cross-language commonalities are stronger at earlier stages of lexical acquisition, compared to later stages. Most of the literature on early word acquisition, however, has been conducted on monolingual children, and neglects the problem of how bilinguals--who represent a substantial proportion of the population in most societies--acquire words at early ages.

@@

Floccia et al.~pointed to \emph{parallel activation} as the main mechanism underpinning their results. The parallel activation hypothesis suggests that bilinguals activate both languages simultaneously during speech production or comprehension. This phenomenon is the result of the activation of lexical representations in both languages, even when only one is in use during production (Costa, Caramazza, \& Sebastian-Galles, 2000; Hoshino \& Kroll, 2008) or comprehension (Thierry \& Wu, 2007). One of the clearest examples of parallel activation was provided by Costa et al. (2000). In this study, Catalan-Spanish monolingual and bilingual adults were asked to name pictures of common objects in Spanish. In half of the trials, the object labels were cognates in Spanish and Catalan (\emph{árbol}-\emph{arbre}, translations of \emph{tree}), whereas in the other half of the trials labels were non-cognates (\emph{mesa}-\emph{taula}, translations of \emph{table}). Bilinguals named cognate pictures faster than non-cognate pictures, even after adjusting for the lexical frequency of the items. Importantly, Spanish monolinguals did not show this effect. These results suggest that, for the bilinguals, Catalan phonology was activated during the production of Spanish words, facilitating the naming of cognate pictures. Several recent studies have also provided similar evidence in comprehension in children (Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013; e.g., Von Holzen \& Mani, 2012). Parallel activation is therefore a plausible mechanism to account for Floccia et al.'s results: cognates increase the amount of activation in both languages, facilitating word acquisition, and ultimately leading to children learning language pairs with a larger proportion of cognates are predicted to show larger vocabulary sizes. A missing piece in this account is how increase parallel activation might facilitate the acquisition new words.

@@

\begin{itemize}
\tightlist
\item
  \emph{Age} of participant in months (\texttt{age\_center}), calculated as the difference in days between participants' birth date and questionnaire completion divided by 30, chunked into bins of two months width, and centred around the mean.
\item
  \emph{Lexical frequency} of item (\texttt{frequency\_center}), Relative lexical frequencies (frequency per million words, \(fpmw\)) were extracted from the CHILDES data base, using the \texttt{childesr} R package (M. Braginsky, Sanchez, \& Yurovsky, 2018), and were transformed to Zipf scores (Van Heuven, Mandera, Keuleers, \& Brysbaert, 2014; Zipf, 1949) (\(log10(fpmw)+3\)). Missing frequencies were imputed by multiple imputation using the \texttt{mice} R package {[}van2011mice{]}. The resulting scores were then centred around the mean.
\item
  \emph{Degree of exposure} to the language (\texttt{doe\_center}), computed as the proportion of the time the child is exposed to the language the item belongs to. For instance, for a response to the item \texttt{perro} (Spanish), the degree of exposure would correspond to the participant's relative exposure to Spanish (ranging from 0 to 1). Before entering the analyses, this variable was centred it around the mean and divided by 10, so that the values of the regression coefficients of this predictor correspond to a 10\% increase in exposure to the language of the item.
\item
  \emph{Cognateness}, a categorical variable indicating whether the Catalan and Spanish forms of the TE are identical cognates (IC), non-identical cognates (NIC), or non-cognates (NC), as rated by a trained Spanish-Catalan bilingual linguist. We sum-coded two contrasts (Schad, Vasishth, Hohenstein, \& Kliegl, 2020): \texttt{cognate1}, comparing IC and NIC (both coded as +0.25) with NC (coded as -0.5) and \texttt{cognate2}, comparing IC (coded as +0.5) with NIC (coded as -0.5).
\end{itemize}

We first fit a base model that only included the main effects of \texttt{age\_center} and \texttt{frequency\_center} (Model 1) (see Mika Braginsky, Yurovsky, Marchman, \& Frank, 2019; Jones \& Brandt, 2019; Kachergis, Marchman, \& Frank, 2021 for similar approaches). We then entered our predictors of interest by adding \texttt{doe\_center} (Model 2) and then \texttt{cognate} and its interaction with \texttt{doe\_center} (Model 3). To account for the hierarchical structure of the data (some responses to different TEs corresponded to the same participant, and some responses from different participants corresponded to the same TE )every model included \emph{Participants} (\texttt{id}) and \emph{TEs} (\texttt{te}) as grouping variables (i.e.~random effects), with all possible random intercepts and slopes where appropriate (Barr, Levy, Scheepers, \& Tily, 2013) (see Appendix for more details of the model).

@@

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{lrrrrrr}
\toprule
& \multicolumn{3}{c}{\textbf{ELPD}} & \multicolumn{3}{c}{\textbf{Diff.}} \\ 
 \cmidrule(lr){2-4}\cmidrule(lr){5-7}
Model & \emph{ELPD} & \emph{SE} & \emph{SE} (sub) & \emph{diff} & \emph{SE} & \emph{SE} (sub) \\ 
\midrule
\multicolumn{1}{l}{Comprehension} \\ 
\midrule
Model 1 & $-233,647.32$ & $688.29$ & $3,175.16$ & - & - & - \\ 
Model 2 & $-241,540.78$ & $716.13$ & $3,477.35$ & $7,893.47$ & $993.27$ & $4,708.89$ \\ 
Model 3 & $-248,415.02$ & $746.19$ & $3,830.96$ & $14,767.71$ & $1,015.16$ & $4,975.74$ \\ 
\midrule
\multicolumn{1}{l}{Production} \\ 
\midrule
Model 1 & $-51,210.58$ & $198.89$ & $106.81$ & - & - & - \\ 
Model 2 & $-55,243.99$ & $204.72$ & $124.16$ & $4,033.42$ & $285.42$ & $163.78$ \\ 
Model 3 & $-60,438.69$ & $207.27$ & $87.25$ & $9,228.11$ & $287.26$ & $137.92$ \\ 
\bottomrule
\end{longtable}

@@

The average probability of participants understanding a TE was 62.37\% (intercept) when all predictors were equal to 0: mean participant age (24.72 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average across the three levels). An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 18.75\% increment in comprehension probability. A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a -0.08\% increment in comprehension probability. A 10\% increase in relative language exposure made TEs -0.10\% more likely to be understood. Cognates (identical and non-identical) were -0.05\% more like to be understood than non-cognates. The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0. Identical cognates were 0.01\% less like to be understood than non-identical cognates. The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero). Figure 1 summarises the posterior distribution of the fixed effects of the model.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{llllrrr}
\toprule
& \multicolumn{3}{c}{Comprehension} & \multicolumn{3}{c}{Production} \\ 
 \cmidrule(lr){2-4}\cmidrule(lr){5-7}
\textbf{Predictor} & \textbf{Mean} & \textbf{SE} & \textbf{95\% CrI} & \textbf{Mean} & \textbf{SE} & \textbf{95\% CrI} \\ 
\midrule
Intercept\textsuperscript{1} & $62.37\%$ & $52.59\%$ & [$57.48\%$, $66.97\%$] & $15.30\%$ & $51.25\%$ & [$14.08\%$, $16.57\%$] \\ 
age\_center\textsuperscript{2} & $18.75\%$ & $2.44\%$ & [$13.79\%$, $23.38\%$] & $5.93\%$ & $0.13\%$ & [$5.66\%$, $6.19\%$] \\ 
frequency\_center\textsuperscript{2} & $-0.08\%$ & $2.54\%$ & [$-5.32\%$, $4.87\%$] & $3.27\%$ & $0.59\%$ & [$2.13\%$, $4.49\%$] \\ 
doe\_center\textsuperscript{2} & $-0.10\%$ & $2.71\%$ & [$-5.41\%$, $5.31\%$] & $14.41\%$ & $0.48\%$ & [$13.49\%$, $15.34\%$] \\ 
cognate1\textsuperscript{2} & $-0.05\%$ & $2.50\%$ & [$-4.99\%$, $4.74\%$] & $2.95\%$ & $2.11\%$ & [$-1.01\%$, $7.16\%$] \\ 
cognate2 & $-0.01\%$ & $2.57\%$ & [$-4.99\%$, $4.95\%$] & $-0.64\%$ & $2.04\%$ & [$-4.67\%$, $3.33\%$] \\ 
doe\_center:cognate1\textsuperscript{2} & $-0.07\%$ & $2.47\%$ & [$-5.15\%$, $4.61\%$] & $-9.05\%$ & $0.83\%$ & [$-10.61\%$, $-7.28\%$] \\ 
doe\_center:cognate2 & $-0.05\%$ & $2.44\%$ & [$-4.85\%$, $4.71\%$] & $-6.53\%$ & $0.73\%$ & [$-7.95\%$, $-5.11\%$] \\ 
\bottomrule
\end{longtable}
\vspace{-5mm}
\begin{minipage}{\linewidth}
\textsuperscript{1}Transformed using the inverse logit to get the average probability of correct response \\ 
\textsuperscript{2}Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable. \\ 
\end{minipage}

@@

The average probability of participants understanding a TE was 15.30\% (intercept) when all predictors were equal to 0: mean participant age (24.72 months), mean item lexical frequency (4.29, Zipf score), mean participant DoE (0.00\%), and regardless of its cognate status (average acros the three levels). An increase in one standard deviation in age (1 \emph{SD} = 3.43 months) was associated with a 5.93\% increment in comprehension probability. A one standard deviation increase in lexical frequency (1 \emph{SD} = 0.71 Zipf scores) was associated with a 3.27\% increment in comprehension probability. A 10\% increase in relative language exposure made TEs 14.41\% more likely to be understood. Cognates (identical and non-identical) were 2.95\% more like to be understood than non-cognates. The 95\% credible interval of the \texttt{doe\_center:cognate1} interaction did not exclude 0. Identical cognates were 0.01\% less like to be understood than non-identical cognates. The 95\% credible interval of the \texttt{doe\_center:cognate2} interaction excluded 0 (although its lower bound was near zero). Figure 1 summarises the posterior distribution of the fixed effects of the model.

\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{lcccrrr}
\toprule
\textbf{Predictor} & \textbf{Mean} & \textbf{SEM} & \textbf{95\% CrI} & \textbf{Rhat} & \textbf{Bulk ESS}\textsuperscript{1} & \textbf{Tail ESS}\textsuperscript{1} \\ 
\midrule
Intercept\textsuperscript{2} & $15.30\%$ & $51.25\%$ & [$14.08\%$, $16.57\%$] & $1.03$ & $108$ & $208$ \\ 
age\_center\textsuperscript{3} & $5.93\%$ & $0.13\%$ & [$5.66\%$, $6.19\%$] & $1.00$ & $620$ & $654$ \\ 
frequency\_center\textsuperscript{3} & $3.27\%$ & $0.59\%$ & [$2.13\%$, $4.49\%$] & $1.01$ & $603$ & $666$ \\ 
doe\_center\textsuperscript{3} & $14.41\%$ & $0.48\%$ & [$13.49\%$, $15.34\%$] & $1.00$ & $675$ & $931$ \\ 
cognate1\textsuperscript{3} & $2.95\%$ & $2.11\%$ & [$-1.01\%$, $7.16\%$] & $1.03$ & $249$ & $419$ \\ 
cognate2 & $-0.64\%$ & $2.04\%$ & [$-4.67\%$, $3.33\%$] & $1.01$ & $307$ & $469$ \\ 
doe\_center:cognate1\textsuperscript{3} & $-9.05\%$ & $0.83\%$ & [$-10.61\%$, $-7.28\%$] & $1.00$ & $1,121$ & $890$ \\ 
doe\_center:cognate2 & $-6.53\%$ & $0.73\%$ & [$-7.95\%$, $-5.11\%$] & $1.00$ & $1,172$ & $790$ \\ 
\bottomrule
\end{longtable}
\vspace{-5mm}
\begin{minipage}{\linewidth}
\textsuperscript{1}ESS: Effective sample size \\ 
\textsuperscript{2}Transformed using the inverse logit to get the average probability of correct response \\ 
\textsuperscript{3}Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable. \\ 
\end{minipage}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-agresti2003categorical}{}}%
Agresti, A. (2003). \emph{Categorical data analysis} (Vol. 482). John Wiley \& Sons.

\leavevmode\vadjust pre{\hypertarget{ref-arslan2020formr}{}}%
Arslan, R. C., Walther, M. P., \& Tata, C. S. (2020). Formr: A study framework allowing for automated feedback generation and complex longitudinal experience-sampling studies using r. \emph{Behavior Research Methods}, \emph{52}(1), 376--387.

\leavevmode\vadjust pre{\hypertarget{ref-barr2013random}{}}%
Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random effects structure for confirmatory hypothesis testing: Keep it maximal. \emph{Journal of Memory and Language}, \emph{68}(3), 255--278.

\leavevmode\vadjust pre{\hypertarget{ref-bergelson2020comprehension}{}}%
Bergelson, E. (2020). The comprehension boost in early word learning: Older infants are better learners. \emph{Child Development Perspectives}, \emph{14}(3), 142--149.

\leavevmode\vadjust pre{\hypertarget{ref-bergelson2012months}{}}%
Bergelson, E., \& Swingley, D. (2012). At 6--9 months, human infants know the meanings of many common nouns. \emph{Proceedings of the National Academy of Sciences}, \emph{109}(9), 3253--3258.

\leavevmode\vadjust pre{\hypertarget{ref-bergelson2015early}{}}%
Bergelson, E., \& Swingley, D. (2015). Early word comprehension in infants: Replication and extension. \emph{Language Learning and Development}, \emph{11}(4), 369--380.

\leavevmode\vadjust pre{\hypertarget{ref-bloom2002children}{}}%
Bloom, P. (2002). \emph{How children learn the meanings of words}. MIT press.

\leavevmode\vadjust pre{\hypertarget{ref-bosch2014first}{}}%
Bosch, L., \& Ramon-Casas, M. (2014). First translation equivalents in bilingual toddlers' expressive vocabulary: {Does} form similarity matter? \emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322. \url{https://doi.org/10.1177/0165025414532559}

\leavevmode\vadjust pre{\hypertarget{ref-braginsky2018childesr}{}}%
Braginsky, M., Sanchez, A., \& Yurovsky, D. (2018). Childesr: Accessing the `CHILDES'database. \emph{R Package Version 0.1. 0}.

\leavevmode\vadjust pre{\hypertarget{ref-braginsky2019consistency}{}}%
Braginsky, Mika, Yurovsky, D., Marchman, V. A., \& Frank, M. C. (2019). Consistency and variability in children's word learning across languages. \emph{Open Mind}, \emph{3}, 52--67.

\leavevmode\vadjust pre{\hypertarget{ref-burkner2017brms}{}}%
Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.

\leavevmode\vadjust pre{\hypertarget{ref-carpenter2017stan}{}}%
Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A. (2017). Stan: A probabilistic programming language. \emph{Journal of Statistical Software}, \emph{76}(1), 1--32.

\leavevmode\vadjust pre{\hypertarget{ref-cattani2014much}{}}%
Cattani, A., Abbot-Smith, K., Farag, R., Krott, A., Arreckx, F., Dennis, I., \& Floccia, C. (2014). How much exposure to english is necessary for a bilingual toddler to perform like a monolingual peer in language tests? \emph{International Journal of Language \& Communication Disorders}, \emph{49}(6), 649--671.

\leavevmode\vadjust pre{\hypertarget{ref-costa2000cognate}{}}%
Costa, A., Caramazza, A., \& Sebastian-Galles, N. (2000). The cognate facilitation effect: Implications for models of lexical access. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{26}(5), 1283.

\leavevmode\vadjust pre{\hypertarget{ref-david2008individual}{}}%
David, A., \& Wei, L. (2008). Individual differences in the lexical development of french--english bilingual children. \emph{International Journal of Bilingual Education and Bilingualism}, \emph{11}(5), 598--618.
